---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# deflist 

<!-- badges: start -->
[![R-CMD-check](https://github.com/bbuchsbaum/deflist/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bbuchsbaum/deflist/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `deflist` is to provide a list-like object for which element access is routed through a user-supplied function. This can be used, for example, when list-like syntax is required but one cannot afford to hold all elements in memory at once.

## Installation

You can install the released version of `deflist` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("deflist")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bbuchsbaum/deflist")
```
## Examples

```{r}

library(deflist)

dl <- deflist(function(i) i, len=10)
print(dl[[1]])

dl2 <- deflist(function(i) { Sys.sleep(5); i*2  }, len=10)

for (i in 1:10) {
  print(dl2[[i]])
}


```

The value at an index may change across calls, for example:

```{r}
dl3 <- deflist(function(i) { rnorm(1) }, len=10)
print(dl3[[1]])
print(dl3[[1]])
```

Memoisation can be enabled so that values at a given index are cached:

```{r}
dl4 <- deflist(function(i) { rnorm(1)  }, len=10, memoise=TRUE)
print(dl4[[1]])
print(dl4[[1]])

```


In addition, memoisation can be set to store cached values to the file system:

```{r}
dl5 <- deflist(function(i) { rnorm(1000)  }, len=10, memoise=TRUE, cache="file", cachedir = tempdir())
print(dl5[[1]][1:10])
print(dl5[[1]][1:10])
```


